Improving \textsl{Gaia} parallax precision \\ with a data-driven model of stars

### 2018-03-04 paper/paper.tex
tweaks to the paper

Although not a fully Bayesian hierarchical inference, the methodology presented in this paper makes exceedingly weak assumptions
about the Milky Way, and similarly weak assumptions about the properties of
stars.
To date, there has been some discussion in the literature
about how to use a parallax responsibly to infer a distance (\citealt{astraatmadja16a}; \citealt{astraatmadja16b}).
These papers use a \gaia\ measurement to construct a parallax likelihood,
and combine this with a distance prior.
While this is sensible and correct, all present methods proposed along
these lines build in informative (and known-to-be-wrong) assumptions
about the line-of-sight distance distribution to \gaia\ stars.
That is, they build in strong and informative assumptions about the Milky
Way.
The strongest new assumption made in this work is that for every star
in \tgas, there are other, photometrically (and bolometrically)
similar stars.
That is a much weaker assumption than has been made in most other Bayesian
distance estimations.

@@

To infer the true parallax for a star using the noise-deconvolved \cmd\ as
a prior, we make the following assumptions. The method we present is correct
and justifiable under these assumptions, each of which is questionable
in its own right. We return to criticize these assumptions---and the
method that flows from them---in the discussion section below.
\begin{description}
\item[stationarity] We make an assumption of \emph{stationarity}; that is, that the
  stars measured at lower signal-to-noise have analogs measured at
  higher signal-to-noise.
  In detail, this assumption is wrong. For example, more distant stars will have lower signal to noise ratio measurements
  and will also be different than local
  disk stars in age and metallicity. However, the stationarity assumption is fairly weak:
  the requirement is that, locally in the \cmd, there is
  support among the high signal-to-noise stars for the types
  of stars also seen at low signal-to-noise.
  If this requirement is met, the results will not be
  strongly biased. Although this is a weak assumption, it is deep and fundamental to this
  work.
\item[selection] We have no model for the selection function of \tgas\ (but see \cite{bovy17}),
  nor selection volumes for any kinds of stars.
  For this reason, we are building a model of \emph{the contents of \tgas}, not the
  properties of the stars in the Milky Way (nor any volume-limited component or region
  thereof). For this reason, our \cmd\ prior might not be appropriate to use with other surveys.
\item[big data] We assume that we have large numbers of stars, large enough that
  empirical-Bayes, or maximum-marginalized-likelihood, is a safe
  approximation to full Bayesian inference. This assumption will be
  least true in the least-populated parts of the \cmd. In the extreme
  case, if a portion of the \cmd\ has only one (statistically distinct)
  exemplar---for example, a white dwarf---inference for these cases will be biased.
\item[noise model] We assume that the \tgas\ parallax uncertainties dominate the
  noise in any estimates of absolute magnitude. We further assume that
  the parallax and color uncertainties are Gaussian in form with
  correctly known variances. In detail, we assume that the parallax uncertainty estimates,
  (the \texttt{parallax{\_}error} entry in the \tgas\ catalog), are correct.
  Here we are making use of the fact that when
  de-noising data, under estimating measurement noise is conservative, and leads to under-deconvolution. In other words, the features of the inferred \cmd\ will typically be broader and more conservatively estimated than if we over-estimated the noise.
\item[dust] We treat the \psone-based three-dimensional dust maps (\citealt{green15})
  as correct in their median dust estimates, and their
  effects on color and magnitude. As we optimize the empirical Bayes model, we iteratively update each star's
  dust correction at its updated, inferred distance. Our assumption is that these corrections are
  good enough, and uncertainties in these are not dominant, for neither
  color nor absolute magnitude.
\item[mixture of Gaussians] We assume that the \cmd\ can be represented by a mixture of
  Gaussians in a particular transformed space. This mixture we fix
  (with only heuristic analysis) at 128 Gaussians.
\item[no physics] We make no use of stellar models; our only assumptions about
  stellar physics are generic and implicit (for example, that the
  \cmd\ is somehow smooth and compact).
\end{description}

@@

Under these assumptions, we would like to infer the true parallax, $\varpi_{\true}$, for each star given its measured parallax, $\varpi$, and associated uncertainty, $\sigma_{\varpi}$, from \tgas.
A first application of Bayes theorem leads to
\begin{equation}
p(\varpi_{\true} \given \varpi, \sigma^2_{\varpi}) = \frac{1}{Z}\,p(\varpi \given \varpi_{\true}, \sigma^2_{\varpi}) \, p(\varpi_{\true}) \quad ,
\label{eq:bayes}
\end{equation}
where $p(\varpi \given \varpi_{\true}, \sigma^2_{\varpi})$ is the likelihood of the observed parallax, $p(\varpi_{\true})$ is a prior PDF for the true
parallax, and $Z$ is a normalization constant (the model evidence, or marginalized likelihood).
The strength of our inference lies in how we generate our prior.
Instead of assuming a functional form for the spatial distribution of stars in the Milky Way (\citealt{astraatmadja16b}), or relying on stellar models (\citealt{gaia16}), we build a noise-deconvolved estimate of the \cmd\ from the observed colors and absolute magnitudes of our full set of stars.
This is because supplementary to the trigonometric measure of distance, the photometry of stars is also informative of their distances.
The intrinsic temperatures and luminosities of stars tightly correlate, leading to the observed colors and absolute magnitudes of stars also tightly correlating, usually visualized as the \cmd.
The tight correlation of the \cmd\ benefits our inference because the luminosity or absolute magnitude of a star $M$, when compared with its brightness or apparent magnitude $m$, contains distance information $d$ via the fundamental relation $m=M+5\log(d/10\, \mathrm{pc})$.
We include this in the inference by using the \cmd\ as part of our prior for $\varpi_{\true}$.
Specifically, we use the \cmd\ generated from the data themselves, as detailed in the next section.

@@

To generate the prior and evaluate the likelihood, we need to
correct the \tmass\ photometry for dust extinction.
With the emergence of 3D dust maps, it is now possible to estimate dust
corrections to stars within the Milky Way. The challenge is getting a
proper estimate of the distance to the star before inferring
$\varpi_{\true}$, since we need a dust correction prior to doing our
inference. Dust corrections are particularly non trivial for stars
with poor parallax signal to noise; these stars have
a likelihood that is consistent with infinite distance and can
therefore have severe dust corrections. This is most obvious in the
giant stars, which tend to have the lowest signal to noise, and can
therefore have dust corrections $> 2$~mag. Although the dust
model is probabilistic, as well as our distance inference, to move
forward, we break from a fully Bayesian framework in which we might
infer the distance and the dust simultaneously (more about this in the
discussion). Instead, we correct for dust using a point estimate from
the 3D dust map \citep{green15}. We determine this dust correction by
iteratively inferring, and then sampling, our posterior PDF over
$\varpi_{\true}$. We first generate our prior using the observed,
attenuated photometry. Using this raw prior, we infer more precise
parallaxes to all the stars in our sample. We use this more
precise parallax posterior to get a measurement of dust in the 3D dust
map for each star. We then iterate this process ten times, however
the dust values seem to converge after a few iterations. The
iterative process allows for slight variations in distance and
corresponding dust to become more consistent with the \cmd.

In detail, we take the $5\%$
quantile of each parallax posterior (the closest part) and query the 3D dust map at that distance.
We take the $5\%$ quantile---as apposed to the median---to do a minimal dust
correction because over-correcting for dust can push the likelihood
function into an unphysical, extremely blue part of the \cmd\ and thus bias the
parallax inference. At the 3D position of each star, we sample the \texttt{mode=median} value of the probabilistic dust map.
We apply each dust correction to our
\tmass\ photometry using Equation~\ref{eq:data}, where $Q_{\lambda}
= [0.709, 0.302]$ for bands $[J, K_s]$ respectively (see Table 6 in
\citealt{schlafly11}).
The dust values we obtain in our final
iteration, and which we apply to the photometry in our inference, are visualized in
galactic coordinates in \figurename~\ref{fig:dust}.
After applying the dust correction, we do not update the uncertainties, which is
technically wrong and under-represents the true uncertainties; this leads to a less severe deconvolution, as mentioned previously.
In other words, the features of our inferred
  \cmd\ are conservatively broad.\footnote{This can be understood by considering the following procedure: if one incrementally increased the noise levels (leaving the data unchanged; just increasing the declared uncertainties), the deconvolved distribution would be increasingly narrower, since it targets the distribution that would generate the noiseless observations, as well as the observed data once noise is added. Thus, under-estimating the uncertainties leads to a conservative under-deconvolution.}
We could add the covariance of
the dust to our likelihoods, but this is beyond the scope of this
demonstration of concept.
\begin{figure}
\centering
  \includegraphics[width=\textwidth]{dust.pdf}
\caption{The converged dust values at the 5\% distance quantile in Galactic coordinates. This shows both the footprint of our analysis as well as the adopted extinction values. Each point represents a single star in our cross-matched catalog. The large missing areas are due to the \psone\ footprint ($>-30~{\rm deg}$ in declination)}
\label{fig:dust}
\end{figure}

@@

@@

We select sources from our catalog within a $1~{\rm deg}^2$ window centered on the position of M67; the cluster is visible as an over-density of stars in the sky positions alone.
To be clear, we do no prior selection on previously known members of M67, but include all stars observed by \tgas\ within this window on the sky.
Keep in mind, the (low) number of M67 members seen here is constrained by the (unknown) selection effects of the \tgas\ dataset.
With our converged dust values, and its associated prior, we calculate the posterior over $\varpi_{\true}$ for each star within this window on the sky around M67.
\figurename~\ref{fig:m67} shows the comparison of the likelihoods and posteriors.
The vertical red lines bracket the previously measured parallaxes of, or distances to, the cluster.
The likelihoods of the observed parallaxes, shown in the top row in black, are broad and the cluster is not at all obvious.
The posterior PDFs, shown in the bottom row in blue, are more sharply peaked showing an increase in precision.
The posterior PDFs also have modes that are more similar to the previously determined distance to the cluster (vertical lines) for far more stars.
This shows that our prior is not only making parallaxes more precise but also (possibly) increasing the accuracy.
\begin{figure}
\centering
  \includegraphics[width=\textwidth]{m67.pdf}
\caption{Proof of concept: parallax inference to the open cluster M67.
The upper left panel visualizes the \gaia\ + \tmass\ likelihoods of the data, shown in black and projected onto the \cmd, for all the stars within our $1~{\rm deg}^2$ window centered on the position of M67.
Our prior, visualized as a mixture of Gaussians, is shown in green.
The upper right panel shows the \tgas\ parallax likelihoods in black, with the ``true" distance to M67, determined by previous methods, bracketed by the red lines.
Similar to the upper left panel, the lower left panel shows the likelihoods of the data in black and our prior in green, as well as the posterior PDFs in blue.
The posterior PDFs are a mixture of Gaussians as well, each a product of the likelihood with a Gaussian in our prior.
The lower right panel shows the projection of the 2D posterior PDF onto parallax space.
The posterior PDFs are much more narrow and clustered than the \tgas\ likelihoods, showing that our prior generates posterior PDFs that are accurate and more precise than \gaia\ alone.}
\label{fig:m67}
\end{figure}

@@

We obtained the photometric parallaxes in this project
by building a data-driven model of \cmd\ of the stars in the \tgas\ data set,
and using it as a prior PDF for Bayesian inference.
The posterior PDFs for distance that we obtain are, in general, much
narrower than the likelihood functions delivered by the
\gaia\ Mission, and therefore the distance estimates (or,
equivalently, parallax estimates) are much more precise.
It is not surprising that a Bayesian inference provides more
precise inferences than the likelihood function alone; Bayesian
inferences bring in new information that decrease variance (but
can introduce bias).
We have shown that, in addition to pure precision improvements, at
least some aspects of \emph{accuracy} have been improved as well:
The posterior distance estimates to stars in (at least one) stellar
cluster are much more clustered than likelihood-based distance
estimates, and they cluster around a sensible value.
Accuracy is demonstrated by this test because each star in the cluster is
subject to a different, unique distance prior (see \figurename~\ref{fig:posterior})
and yet the posterior PDFs are all pulled to the same parallax value.

@@

\subsection{Comparison with other priors}
\begin{figure}
\centering
  \includegraphics[width=\textwidth]{comparePrior.pdf}
\caption{Comparison with a common spatial prior: A visualization of the posterior PDF for a subsection of \tgas\ stars, shown as the expectation value and $1\sigma$ uncertainty. The left panel shows the posterior PDF using the exponentially declining spatial density, a common space-density prior used when inferring distances. The right panel shows the posterior PDF using our \cmd\ prior. The scale of the exponentially declining spatial density prior is $1.35~{\rm kpc}$, the optimal value found in \cite{astraatmadja16a}. Using the \cmd\ as a prior generates a posterior PDF that has smaller variance.}
\label{fig:comparePrior}
\end{figure}

There are other options for prior information to include when inferring distances to stars.
One option is the exponentially declining spatial density (EDSD) of stars \citep{astraatmadja16a}.
It is parameterized by a scale length, and has the nice property that it is smooth out to very large distances so that the posteriors are also very smooth.
It is a fairly weak prior.
\figurename~\ref{fig:comparePrior} shows a comparison of the EDSD prior with our \cmd\ prior.
Here the parallax posterior PDF expectation value, and $1\sigma$ uncertainty, for a subset of stars in our dataset is converted into an absolute magnitude and visualized as the \cmd.
The left panel visualizes results using the EDSD prior, and the right panel visualizes results using our \cmd\ prior.
The variance for the \cmd\ prior is significantly smaller than the variance for the EDSD prior.
In general, EDSD is a fairly broad prior so the median \gaia\ star has a posterior variance equal to the likelihood variance.
EDSD has the largest effect on the most noisy and negative parallaxes, and has the nice property of bringing these noisy or negative measurements to reasonable parallaxes.

@@

  prior to infer distances for the full billion-star catalog released

@@

  at a quantile of our distance posterior, correcting our photometry

>>>>

Although not a fully Bayesian hierarchical inference, the methodology presented in this paper makes exceedingly weak assumptions
about the Milky Way, and similarly weak assumptions about the properties of
stars.
To date, there has been some discussion in the literature
about how to use a parallax responsibly to infer a distance (\citealt{astraatmadja16a}; \citealt{astraatmadja16b}).
These papers use a \gaia\ measurement to construct a parallax likelihood,
and combine this with a parallax prior.
While this is sensible and correct, all present methods proposed along
these lines build in informative (and known-to-be-wrong) assumptions
about the line-of-sight parallax distribution to \gaia\ stars.
That is, they build in strong and informative assumptions about the Milky
Way.
The strongest new assumption made in this work is that for every star
in \tgas, there are other, photometrically (and bolometrically)
similar stars.
That is a much weaker assumption than has been made in most other Bayesian
distance estimations.

@@

To infer the true parallax for a star using the noise-deconvolved \cmd\ as
a prior, we make the following assumptions. The method we present is correct
and justifiable under these assumptions, each of which is questionable
in its own right. We return to criticize these assumptions---and the
method that flows from them---in the discussion section below.
\begin{description}
\item[stationarity] We make an assumption of \emph{stationarity}; that is, that the
  stars measured at lower signal-to-noise have analogs measured at
  higher signal-to-noise.
  In detail, this assumption is wrong. For example, more distant stars will have lower signal to noise ratio measurements
  and will also be different than local
  disk stars in age and metallicity. However, the stationarity assumption is fairly weak:
  the requirement is that, locally in the \cmd, there is
  support among the high signal-to-noise stars for the types
  of stars also seen at low signal-to-noise.
  If this requirement is met, the results will not be
  strongly biased. Although this is a weak assumption, it is deep and fundamental to this
  work.
\item[selection] We have no model for the selection function of \tgas\ (but see \cite{bovy17}),
  nor selection volumes for any kinds of stars.
  For this reason, we are building a model of \emph{the contents of \tgas}, not the
  properties of the stars in the Milky Way (nor any volume-limited component or region
  thereof). For this reason, our \cmd\ prior might not be appropriate to use with other surveys.
\item[big data] We assume that we have large numbers of stars, large enough that
  empirical-Bayes, or maximum-marginalized-likelihood, is a safe
  approximation to full Bayesian inference. This assumption will be
  least true in the least-populated parts of the \cmd. In the extreme
  case, if a portion of the \cmd\ has only one (statistically distinct)
  exemplar---for example, a white dwarf---inference for these cases will be biased.
\item[noise model] We assume that the \tgas\ parallax uncertainties dominate the
  noise in any estimates of absolute magnitude. We further assume that
  the parallax and color uncertainties are Gaussian in form with
  correctly known variances. In detail, we assume that the parallax uncertainty estimates,
  (the \texttt{parallax{\_}error} entry in the \tgas\ catalog), are correct.
  Here we are making use of the fact that when
  de-noising data, under estimating measurement noise is conservative, and leads to under-deconvolution. In other words, the features of the inferred \cmd\ will typically be broader and more conservatively estimated than if we over-estimated the noise.
\item[dust] We treat the \psone-based three-dimensional dust maps (\citealt{green15})
  as correct in their median dust estimates, and their
  effects on color and magnitude. As we optimize the empirical Bayes model, we iteratively update each star's
  dust correction at its updated, inferred parallax. Our assumption is that these corrections are
  good enough, and uncertainties in these are not dominant, for neither
  color nor absolute magnitude.
\item[mixture of Gaussians] We assume that the \cmd\ can be represented by a mixture of
  Gaussians in a particular transformed space. This mixture we fix
  (with only heuristic analysis) at 128 Gaussians.
\item[no physics] We make no use of stellar models; our only assumptions about
  stellar physics are generic and implicit (for example, that the
  \cmd\ is somehow smooth and compact).
\end{description}

@@

Under these assumptions, we would like to infer the true parallax, $\varpi_{\true}$, for each star given its measured parallax, $\varpi$, and associated uncertainty, $\sigma_{\varpi}$, from \tgas.
A first application of Bayes theorem leads to
\begin{equation}
p(\varpi_{\true} \given \varpi, \sigma^2_{\varpi}) = \frac{1}{Z}\,p(\varpi \given \varpi_{\true}, \sigma^2_{\varpi}) \, p(\varpi_{\true}) \quad ,
\label{eq:bayes}
\end{equation}
where $p(\varpi \given \varpi_{\true}, \sigma^2_{\varpi})$ is the likelihood of the observed parallax, $p(\varpi_{\true})$ is a prior PDF for the true
parallax, and $Z$ is a normalization constant (the model evidence, or marginalized likelihood).
The strength of our inference lies in how we generate our prior.
Instead of assuming a functional form for the spatial distribution of stars in the Milky Way (\citealt{astraatmadja16b}), or relying on stellar models (\citealt{gaia16}), we build a noise-deconvolved estimate of the \cmd\ from the observed colors and absolute magnitudes of our full set of stars.
This is because supplementary to the trigonometric measure of distance, the photometry of stars is also informative of their distances.
The intrinsic temperatures and luminosities of stars tightly correlate, leading to the observed colors and absolute magnitudes of stars also tightly correlating, usually visualized as the \cmd.
The tight correlation of the \cmd\ benefits our inference because the luminosity or absolute magnitude of a star $M$, when compared with its brightness or apparent magnitude $m$, contains \added{parallax information $\varpi$ via the fundamental relation $m=M+10-5\log(\varpi[\mathrm{mas}])$.}
We include this in the inference by using the \cmd\ as part of our prior for $\varpi_{\true}$.
Specifically, we use the \cmd\ generated from the data themselves, as detailed in the next section.

@@

To generate the prior and evaluate the likelihood, we need to
correct the \tmass\ photometry for dust extinction.
With the emergence of 3D dust maps, it is now possible to estimate dust
corrections to stars within the Milky Way. The challenge is getting a
proper estimate of the distance to the star before inferring
$\varpi_{\true}$, since we need a dust correction prior to doing our
inference. Dust corrections are particularly non trivial for stars
with poor parallax signal to noise; these stars have
a likelihood that is consistent with infinite distance and can
therefore have severe dust corrections. This is most obvious in the
giant stars, which tend to have the lowest signal to noise, and can
therefore have dust corrections $> 2$~mag. Although the dust
model is probabilistic, as well as our parallax inference, to move
forward, we break from a fully Bayesian framework in which we might
infer the parallax and the dust simultaneously (more about this in the
discussion). Instead, we correct for dust using a point estimate from
the 3D dust map \citep{green15}. We determine this dust correction by
iteratively inferring, and then sampling, our posterior PDF over
$\varpi_{\true}$. We first generate our prior using the observed,
attenuated photometry. Using this raw prior, we infer more precise
parallaxes to all the stars in our sample. We use this more
precise parallax posterior to get a measurement of dust in the 3D dust
map for each star. We then iterate this process ten times, however
the dust values seem to converge after a few iterations. The
iterative process allows for slight variations in parallax and
corresponding dust to become more consistent with the \cmd.

In detail, we take the $5\%$
quantile of each parallax posterior (the closest part) and query the 3D dust map at that distance.
We take the $5\%$ quantile---as apposed to the median---to do a minimal dust
correction because over-correcting for dust can push the likelihood
function into an unphysical, extremely blue part of the \cmd\ and thus bias the
parallax inference. At the 3D position of each star, we sample the \texttt{mode=median} value of the probabilistic dust map.
We apply each dust correction to our
\tmass\ photometry using Equation~\ref{eq:data}, where $Q_{\lambda}
= [0.709, 0.302]$ for bands $[J, K_s]$ respectively (see Table 6 in
\citealt{schlafly11}).
The dust values we obtain in our final
iteration, and which we apply to the photometry in our inference, are visualized in
galactic coordinates in \figurename~\ref{fig:dust}.
After applying the dust correction, we do not update the uncertainties, which is
technically wrong and under-represents the true uncertainties; this leads to a less severe deconvolution, as mentioned previously.
In other words, the features of our inferred
  \cmd\ are conservatively broad.\footnote{This can be understood by considering the following procedure: if one incrementally increased the noise levels (leaving the data unchanged; just increasing the declared uncertainties), the deconvolved distribution would be increasingly narrower, since it targets the distribution that would generate the noiseless observations, as well as the observed data once noise is added. Thus, under-estimating the uncertainties leads to a conservative under-deconvolution.}
We could add the covariance of
the dust to our likelihoods, but this is beyond the scope of this
demonstration of concept.
\begin{figure}
\centering
  \includegraphics[width=\textwidth]{dust.pdf}
\caption{The converged dust values at the 5\% parallax quantile in Galactic coordinates. This shows both the footprint of our analysis as well as the adopted extinction values. Each point represents a single star in our cross-matched catalog. The large missing areas are due to the \psone\ footprint ($>-30~{\rm deg}$ in declination)}
\label{fig:dust}
\end{figure}

@@

\begin{figure}
\centering
  \includegraphics[width=\textwidth]{posterior_odd.pdf}
\caption{Five example posterior PDFs over parallax using our method for more odd objects.
The upper left panel shows the position of each example data point in black, with its observed uncertainties, plotted over our prior represented by a mixture of Gaussians in green. In blue, we show the posterior expectation value and $1 \sigma$ uncertainties. 
We chose example points that lie outside the well populated portions of the \cmd to illustrate how our method works for rare \tgas objects. 
The remaining panels show the posterior PDF in blue, the prior PDF in green, and likelihood of the data in black.

Most stars have posterior PDFs (blue) that are only slightly more precise than the data (black), which is well represented in this random sampling of points.
Examples 1, 3, and 4 are not very effected by the prior (green), where good data is good data.
Examples 2 and 5 show a more precise posterior PDF compared with the likelihood.
Example 5 looks very similar to the prior PDF and exemplifies that our prior tends to have the largest effect on stars in the red giant branch.}
\label{fig:posterior}
\end{figure}

@@

We select sources from our catalog within a $0.75~{\rm deg}^2$ window centered on the position of M67; the cluster is visible as an over-density of stars in the sky positions alone.
To be clear, we do no prior selection on previously known members of M67, but include all stars observed by \tgas\ within this window on the sky.
Keep in mind, the (low) number of M67 members seen here is constrained by the (unknown) selection effects of the \tgas\ dataset.
With our converged dust values, and its associated prior, we calculate the posterior over $\varpi_{\true}$ for each star within this window on the sky around M67.
\figurename~\ref{fig:m67} shows the comparison of the likelihoods and posteriors.
The vertical red dotted lines bracket the previously measured parallaxes of, or distances to, the cluster.
The likelihoods of the observed parallaxes, shown in the top row in black, are broad and the cluster is not at all obvious.
The posterior PDFs, shown in the bottom row in blue, are more sharply peaked showing an increase in precision.
The posterior PDFs also have modes that are more similar to the previously determined parallax of the cluster (vertical lines) for far more stars.
This shows that our prior is not only making parallaxes more precise but also (possibly) increasing the accuracy.
\begin{figure}
\centering
  \includegraphics[width=\textwidth]{m67.pdf}
\caption{Proof of concept: parallax inference to the open cluster M67.
The upper left panel visualizes the \gaia\ + \tmass\ likelihoods of the data, shown in black and projected onto the \cmd, for all the stars within our $0.75~{\rm deg}^2$ window centered on the position of M67.
Our prior, visualized as a mixture of Gaussians, is shown in green.
The upper right panel shows the \tgas\ parallax likelihoods in black, with the ``true" parallax to M67, determined by previous methods, bracketed by the red dotted lines.
Similar to the upper left panel, the lower left panel shows the likelihoods of the data in black and our prior in green, as well as the posterior PDFs in blue.
The posterior PDFs are a mixture of Gaussians as well, each a product of the likelihood with a Gaussian in our prior.
The lower right panel shows the projection of the 2D posterior PDF onto parallax space.
The posterior PDFs are much more narrow and clustered than the \tgas\ likelihoods, showing that our prior generates posterior PDFs that are accurate and more precise than \gaia\ alone.}
\label{fig:m67}
\end{figure}

@@

We obtained the photometric parallaxes in this project
by building a data-driven model of \cmd\ of the stars in the \tgas\ data set,
and using it as a prior PDF for Bayesian inference.
The posterior PDFs for parallax that we obtain are, in general, much
narrower than the likelihood functions delivered by the
\gaia\ Mission, and therefore the distance estimates (or,
equivalently, parallax estimates) are much more precise.
It is not surprising that a Bayesian inference provides more
precise inferences than the likelihood function alone; Bayesian
inferences bring in new information that decrease variance (but
can introduce bias).
We have shown that, in addition to pure precision improvements, at
least some aspects of \emph{accuracy} have been improved as well:
The posterior parallax estimates to stars in (at least one) stellar
cluster are much more clustered than likelihood-based parallax
estimates, and they cluster around a sensible value.
Accuracy is demonstrated by this test because each star in the cluster is
subject to a different, unique parallax prior (see \figurename~\ref{fig:posterior})
and yet the posterior PDFs are all pulled to the same parallax value.

@@

\subsection{Comparison with other priors}
\begin{figure}
\centering
  \includegraphics[width=\textwidth]{comparePrior.pdf}
\caption{Comparison with a common spatial prior: A visualization of the posterior PDF for a subsection of \tgas\ stars, shown as the expectation value and $1\sigma$ uncertainty. The left panel shows the posterior PDF using the exponentially declining spatial density, a common space-density prior used when inferring parallax. The right panel shows the posterior PDF using our \cmd\ prior. The scale of the exponentially declining spatial density prior is $1.35~{\rm kpc}$, the optimal value found in \cite{astraatmadja16a}. Using the \cmd\ as a prior generates a posterior PDF that has smaller variance.}
\label{fig:comparePrior}
\end{figure}

There are other options for prior information to include when inferring the parallax of a star.
One option is the exponentially declining spatial density (EDSD) of stars \citep{astraatmadja16a}.
It is parameterized by a scale length, and has the nice property that it is smooth out to very large distances so that the posteriors are also very smooth.
It is a fairly weak prior.
\figurename~\ref{fig:comparePrior} shows a comparison of the EDSD prior with our \cmd\ prior.
Here the parallax posterior PDF expectation value, and $1\sigma$ uncertainty, for a subset of stars in our dataset is converted into an absolute magnitude and visualized as the \cmd.
The left panel visualizes results using the EDSD prior, and the right panel visualizes results using our \cmd\ prior.
The variance for the \cmd\ prior is significantly smaller than the variance for the EDSD prior.
In general, EDSD is a fairly broad prior so the median \gaia\ star has a posterior variance equal to the likelihood variance.
EDSD has the largest effect on the most noisy and negative parallaxes, and has the nice property of bringing these noisy or negative measurements to reasonable parallaxes.

@@

  prior to infer parallaxes for the full billion-star catalog released

@@

  at a quantile of our parallax posterior, correcting our photometry
